---
title: "A simple R Notebook for exploring 2012 time series data"
output:
  pdf_document: default
  html_notebook: default
---
#install the packages
```{r load libraries, warning=FALSE, message=FALSE, include=FALSE}
packages.used=as.list(
  c(
  "tidyverse",
  "haven",
  "devtools",
  "RColorBrewer",
  "data.table",
  "ggplot2")
)

check.pkg = function(x){
  if(!require(x, character.only=T)) install.packages(x, 
                                                     character.only=T,
                                                     dependence=T)
}

lapply(packages.used, check.pkg)
```


#import the data
```{r read in data, message = F}
library(haven)
anes_timeseries_2012 <-
    read_sav("~/Documents/GitHub/Spring2021-Project1-Yuqi-Xing/data/anes_timeseries_2012.sav")
```

### Process variables for analysis
```{r labelled variables subset}
anes_new=anes_timeseries_2012%>%
  mutate(
    turnout=as_factor(rvote2012_x),
    vote_for_president=as_factor(presvote2012_x),
    vote_for_USHouse=as_factor(hsevote2012_x),
    vote_for_senvote=as_factor(senvote2012_x),
    race=as_factor(dem_raceeth_x),
    gender=as_factor(gender_respondent_x),
  )

library(data.table)

data.table(anes_new%>%
             select(turnout, vote_for_president, vote_for_USHouse, vote_for_senvote, race, gender)%>%
             filter(turnout!="-9"&turnout!="-6"&turnout!="-2. Missing, 2012 voting status not determined")%>%
             sample_n(10))
anes_new = anes_new%>%
  filter(turnout!="-9"&turnout!="-6"&turnout!="-2. Missing, 2012 voting status not determined")%>%
  select(turnout, vote_for_president, vote_for_USHouse, vote_for_senvote, race, gender)
head(anes_new)
save(anes_new, file="~/Documents/GitHub/Spring2021-Project1-Yuqi-Xing/output/data_use.RData")
```
#  A simple analysis
#Who went to vote in the election?
```{r turnout analysis}
load(file="~/Documents/GitHub/Spring2021-Project1-Yuqi-Xing/output/data_use.RData")
anes_to_race = anes_new %>%
  filter(race!="-9. Missing")     ###delete useless data from race variable

anes_to_gender = anes_new
  
ggplot(anes_to_race)+
  geom_bar(aes(x=turnout, fill=race)) +
  theme(axis.text.x = element_text(angle = 50, hjust = 1))+
  labs(title="How did different racial groups participate in the election?")+
  scale_x_discrete(labels=c("1. R voted in the 2012 elections" = "voted", "2. R did not vote in the 2012 elections" = "did not vote"))+
  scale_fill_discrete(name="race", breaks=c("1. White, non-Hispanic", "2. Black, non-Hispanic", "3. Asian, native Hawaiian or other Pacif Islr, non-Hispanic","4. Native American or Alaska Native, non-Hispanic","5. Hispanic", "6. Other non-Hispanic incl multiple races (Web: blank 'Other' counted as a race)"), labels=c("1. White,non-Hispanic", "2. Black,non-Hispanic", "3. Asian,native Hawaiian \n or other Pacif Islr, \n non-Hispanic", "4. Native American or \n Alaska Native, \n non-Hispanic", "5. Hispanic", "6. Other non-Hispanic \n incl multiple races"))   ##create bar graph for turnout, according to different race

ggplot(anes_to_race)+
  geom_bar(aes(x=race, fill=turnout)) +
  theme(axis.text.x = element_text(angle = 50, hjust = 1))+
labs(title="How did different racial groups participate in the election?")+
scale_x_discrete(labels=c("1. White,non-Hispanic", "2. Black,non-Hispanic", "3. Asian,native Hawaiian \n or other Pacif Islr, \n non-Hispanic", "4. Native American or \n Alaska Native, \n non-Hispanic", "5. Hispanic", "6. Other non-Hispanic \n incl multiple races"))+
  scale_fill_discrete(name="race", breaks=c("1. R voted in the 2012 elections", "2. R did not vote in the 2012 elections"), labels=c("voted", "did not vote"))    #create bar graph for race, according to different turnout 

ggplot(anes_to_gender)+
  geom_bar(aes(x=turnout, fill=gender)) +
  theme(axis.text.x = element_text(angle = 50, hjust = 1, size=7))+
  labs(title="How did different gender groups participate in the election?")+
  scale_x_discrete(labels=c("1. R voted in the 2012 elections" = "voted", "2. R did not vote in the 2012 elections" = "did not vote"))     # create bar graph for turnout, according to different gender 
```
conclusion:
Overall, most people turned out to vote.
For each race, the number of people who voted was significantly greater than the number who did not vote.
Because of the large white population in the United States, whites play an important role in voting, like whites voting more than all other races combining.
There is no obvious gender bias in voting participation.


#Who did they vote for as president in the election?
```{r vote for presient analysis}
anes_vote_for_president=anes_new %>%
  filter(vote_for_president!="-9"&vote_for_president!="-6"&vote_for_president!="-2. R did not vote for Pres or did not report vote for Pres in pre and no post-election data")   # delete the useless data for vote_for_president variable

anes_vote_for_president_by_race = anes_vote_for_president%>%
  filter(race!="-9. Missing")    # delete the useless race data for vote_for_president variable


anes_vote_for_president_by_gender= anes_vote_for_president

ggplot(anes_vote_for_president_by_race)+
geom_bar(aes(x=vote_for_president, fill=race))+
labs(title="Who did different racial groups vote for president in the election?")+
theme(axis.text.x = element_text(angle = 45, hjust = 1))+
scale_fill_discrete(name="race", breaks=c("1. White, non-Hispanic", "2. Black, non-Hispanic", "3. Asian, native Hawaiian or other Pacif Islr, non-Hispanic","4. Native American or Alaska Native, non-Hispanic","5. Hispanic", "6. Other non-Hispanic incl multiple races (Web: blank 'Other' counted as a race)"), labels=c("1. White,non-Hispanic", "2. Black,non-Hispanic", "3. Asian,native Hawaiian \n or other Pacif Islr, \n non-Hispanic", "4. Native American or \n Alaska Native, \n non-Hispanic", "5. Hispanic", "6. Other non-Hispanic \n incl multiple races"))  ##create bar graph for vote_for_president, according to different race

ggplot(anes_vote_for_president_by_race)+
geom_bar(aes(x=race, fill=vote_for_president))+
labs(title="Who did different racial groups vote for president in the election?")+
theme(axis.text.x = element_text(angle = 45, hjust = 1))+
labs(title="Who did different racial groups vote for president in the election?")+
scale_x_discrete(labels=c("1. White,non-Hispanic", "2. Black,non-Hispanic", "3. Asian,native Hawaiian \n or other Pacif Islr, \n non-Hispanic", "4. Native American or \n Alaska Native, \n non-Hispanic", "5. Hispanic", "6. Other non-Hispanic \n incl multiple races"))   #create bar graph for race, according to different vote_for_president

ggplot(anes_vote_for_president_by_gender)+
geom_bar(aes(x=vote_for_president, fill=gender))+
labs(title="Who did different gender groups vote for president in the election?")+
theme(axis.text.x = element_text(angle = 45, hjust = 1))   ##create bar graph for vote_for_president, according to different gender
```
conclusion:
Mr Obama scored significantly higher than the other two.
Among whites, nearly half support Barack Obama, but more than half support Mitt Romney.
Among blacks, nearly all voted for Obama (Obanma is black, so there may exist racial bias).
Among Hispanic, most people voted for Obanma.
For the other races, the vote was evenly split, or close to evenly split by Obama and Romney.
There is no obvious gender bias in voting for president.

#which party did they vote for as U.S. House in the election?
```{r vote for U.S. House analysis}
anes_vote_for_USHouse=anes_new %>%
  filter(vote_for_USHouse!="-1. Inap, R did not or DK/RF if voted; voted but not (or DK/RF if voted for) for U.S. House"&vote_for_USHouse!="-6. No post-election data (deleted partial or no post IW) and R did not vote in pre-election (early vote)"&vote_for_USHouse!="-2. Missing, DK/RF for whom R voted")   # delete the useless data for vote_for_USHouse variable

anes_vote_for_USHouse_by_race = anes_vote_for_USHouse%>%
  filter(race!="-9. Missing")   # delete the useless race data for vote_for_USHouse variable


anes_vote_for_USHouse_by_gender= anes_vote_for_USHouse

ggplot(anes_vote_for_USHouse_by_race)+
geom_bar(aes(x=vote_for_USHouse, fill=race))+
theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_x_discrete(labels=c("1. Voted for Democratic House candidate" = "Democratic \n House candidate", "2. Voted for Republican House candidate" = "Republican \n House candidate", "7. Voted for other House candidate" = "other House \n candidate"))+
  labs(title="Which party did different racial groups vote for as U.S. House in the election?")+
  scale_fill_discrete(name="race", breaks=c("1. White, non-Hispanic", "2. Black, non-Hispanic", "3. Asian, native Hawaiian or other Pacif Islr, non-Hispanic","4. Native American or Alaska Native, non-Hispanic","5. Hispanic", "6. Other non-Hispanic incl multiple races (Web: blank 'Other' counted as a race)"), labels=c("1. White,non-Hispanic", "2. Black,non-Hispanic", "3. Asian,native Hawaiian \n or other Pacif Islr, \n non-Hispanic", "4. Native American or \n Alaska Native, \n non-Hispanic", "5. Hispanic", "6. Other non-Hispanic \n incl multiple races"))   ##create bar graph for vote_for_USHouse, according to different race

ggplot(anes_vote_for_USHouse_by_race)+
geom_bar(aes(x=race, fill=vote_for_USHouse))+
theme(axis.text.x = element_text(angle = 45, hjust = 1))+
labs(title="Which party did different racial groups vote for as U.S. House in the election?")+
scale_x_discrete(labels=c("1. White,non-Hispanic", "2. Black,non-Hispanic", "3. Asian,native Hawaiian \n or other Pacif Islr, \n non-Hispanic", "4. Native American or \n Alaska Native, \n non-Hispanic", "5. Hispanic", "6. Other non-Hispanic \n incl multiple races"))+
  scale_fill_discrete(name="vote for U.S. House", breaks=c("1. Voted for Democratic House candidate", "2. Voted for Republican House candidate", "7. Voted for other House candidate"), labels=c("1. Democratic House candidate", "2. Republican House candidate", "7. other House candidate"))    ##create bar graph for race, according to different vote_for_USHouse


ggplot(anes_vote_for_USHouse_by_gender)+
geom_bar(aes(x=vote_for_USHouse, fill=gender))+
theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  labs(title="Which party did different gender groups vote for as U.S. House in the election?")+
  scale_x_discrete(labels=c("1. Voted for Democratic House candidate" = "Democratic \n House candidate", "2. Voted for Republican House candidate" = "Republican \n House candidate", "7. Voted for other House candidate" = "other House \n candidate"))   ##create bar graph for vote_for_USHouse, according to different gender

```
conclusion:
Overall, the majority favored Democrats, followed by Republicans.
Most whites supported the Republicans.
Among blacks, nearly all voted for Democrats (Obanma is black & Democrats, so there may exist racial bias).
Among Hispanic, most people voted for Democrats.
For the other races, the vote was evenly split, or close to evenly split by Democats and Republican.
There is no obvious gender bias in voting for USHouse.


#which party did they vote for as Senvote in the election?
```{r vote for senvote analysis}
anes_vote_for_senvote=anes_new %>%
  filter(vote_for_senvote!="-1. Inap, R did not vote or DK/RF if voted; voted but not (or DK/RF if) for us Senate; no us Sen race in state of vote"&vote_for_senvote!="-6. No post-election data (deleted partial or no post IW) and R did not vote in pre-election (early vote)"&vote_for_senvote!="-2. Missing, DK/RF for whom R voted")     # delete the useless data for vote_for_senvote variable

anes_vote_for_senvote_by_race = anes_vote_for_senvote%>%
  filter(race!="-9. Missing")    # delete the useless race data for vote_for_senvote variable


anes_vote_for_senvote_by_gender= anes_vote_for_senvote

ggplot(anes_vote_for_senvote_by_race)+
geom_bar(aes(x=vote_for_senvote, fill=race))+
theme(axis.text.x = element_text(angle = 45, hjust = 1))+
labs(title="Which party did different racial groups vote for as Senvote in the election?")+
scale_x_discrete(labels=c("1. Voted for Democratic Senate candidate" = "Democratic \n Senate candidate", "2. Voted for Republican Senate candidate" = "Republican \n Senate candidate", "7. Voted for other Senate candidate" = "other Senate \n candidate"))+
  scale_fill_discrete(name="race", breaks=c("1. White, non-Hispanic", "2. Black, non-Hispanic", "3. Asian, native Hawaiian or other Pacif Islr, non-Hispanic","4. Native American or Alaska Native, non-Hispanic","5. Hispanic", "6. Other non-Hispanic incl multiple races (Web: blank 'Other' counted as a race)"), labels=c("1. White,non-Hispanic", "2. Black,non-Hispanic", "3. Asian,native Hawaiian \n or other Pacif Islr, \n non-Hispanic", "4. Native American or \n Alaska Native, \n non-Hispanic", "5. Hispanic", "6. Other non-Hispanic \n incl multiple races"))    ##create bar graph for vote_for_senvote, according to different race

ggplot(anes_vote_for_senvote_by_race)+
geom_bar(aes(x=race, fill=vote_for_senvote))+
theme(axis.text.x = element_text(angle = 45, hjust = 1))+
labs(title="Which party did different racial groups vote for as Senvote in the election?")+
scale_x_discrete(labels=c("1. White,non-Hispanic", "2. Black,non-Hispanic", "3. Asian,native Hawaiian \n or other Pacif Islr, \n non-Hispanic", "4. Native American or \n Alaska Native, \n non-Hispanic", "5. Hispanic", "6. Other non-Hispanic \n incl multiple races"))+
  scale_fill_discrete(name="vote for Senvote", breaks=c("1. Voted for Democratic Senate candidate", "2. Voted for Republican Senate candidate", "7. Voted for other Senate candidate"), labels=c("1. Democratic Senate candidate", "2. Republican Senate candidate", "7. other Senate candidate"))     ##create bar graph for race, according to different vote_for_Senate


ggplot(anes_vote_for_senvote_by_gender)+
geom_bar(aes(x=vote_for_senvote, fill=gender))+
theme(axis.text.x = element_text(angle = 45, hjust = 1))+
labs(title="Which party did different racial groups vote for as Senvote in the election?")+
scale_x_discrete(labels=c("1. Voted for Democratic Senate candidate" = "Democratic \n Senate candidate", "2. Voted for Republican Senate candidate" = "Republican \n Senate candidate", "7. Voted for other Senate candidate" = "other Senate \n candidate"))    ##create bar graph for vote_for_Senvote, according to different gender
```
conclusion:
Overall, the majority favored Democrats, followed by Republicans.
Half of whites voted Democratic and half of whites supported the Republicans.
Among blacks, nearly all voted for Democrats (Obanma is black & Democrats, so there may exist racial bias).
Among Hispanic, most people voted for Democrats.
For the other races, the vote was evenly split, or close to evenly split by Democats and Republican.
There is no obvious gender bias in voting for Senate.